Algorithm Engineering

### 2022-11-30 Assignments/paper.tex
06

Atomicity means that something can’t be divided into smaller pieces. Threads need to use atomic reads and writes to share data. 
If they are not atomic they are so-called torn reads and writes.
To ensure atomicity one need to make sure that variables used for thread synchronization aren’t larger than the CPU word size.
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{8}

>>>>

Atomicity means that something can’t be divided into smaller pieces. Threads need to use atomic reads and writes to share data. 
If they are not atomic they are so-called torn reads and writes.
To ensure atomicity one need to make sure that variables used for thread synchronization aren’t larger than the CPU word size.

% (5)
\section{Assignment}

\subsection{What is CMake?}
CMake is a script language used to create build files that can be executed across different platforms.
The build files are executed using a compiler. CMake provides tools to build, test and package software.

\subsection{What role do targets play in CMake?}
Targets define what is build. They are executables and libraries.
Targets have constructors to build executables, libraries or tests.
Targets can be thought of as objects with different properties. To change these properties we can use member functions.
Within member functions the properties can be specified. Properties can flags, directories or linked libraries.

\subsection{How would you proceed to optimize code?}
The most important thing is to try solving the problem. Once that is done and my code works, I would look at the performance.
If performance can be improved, I would look at ways to do so. I would try parallelizing sections of code, try finding ways to improve the algorithm.
I would continue doing so until the performance is good.

% (6)
\section{Assignment}

\subsection{Name some characteristics of the instruction sets: SSE, AVX(2), AVX-512}
SSE, AVX2, AVX-512 refer to instruction sets on Intel CPUs. They differ in the supported vector length and the number of registers.

SSE: 128-bit vector length, 1999 - 2009 years of launch, 16 registers
AVX2: 256-bit vector length, 2011 / 2013 years of launch, 16 registers
AVX-512: 512-bit vector length, 2017 year of launch, 32 registers

CPUs for high-performance computing support AVX-512 instruction set, wehere as modern mainstream CPUs support AVX2 at most.

\subsection{How can memory aliasing affect performance?}
Memory aliasing occurs when two pointers point to the same memory location.
The compiler doesn't know if two pointers might point to the same address. 
To ensure safe optimizations, the compiler always assumes pointer aliasing. 
Therefore, we can give the compiler hints, to ensure it that no memory aliasing was made.
This can be done using the restrict keyword (\*\_\_restrict\_\_ for gcc). Memory aliasing is important for pointers as well as references.

\subsection{What are the advantages of unit stride (stride-1) memory access compared to accessing memory with larger strides (for example stride-8)?}
stride-1 - access sequential elements in memory, where each fields are equally distant. The distance in this case is called stride.
The difference between stride-1 and larger strides stride-n is that every nth element is accessed.

Stride-1 is faster than larger strides. There is a maximum bandwith that can be loaded from memory. 
Using larger strides leads to loading a higher bandwidth without fully using it.

We can check this by looking at the summation of two arrays. 
Summing up every 8th element is more or less as performant as summing up every element of the array.

\subsection{When would you prefer arranging records in memory as a Structure of Arrays?}
There are two ways to arrange records in memory: as Array of Structures or Structure of Arrays.
Structure of Arrays: one struct with multiple arrays.
Array of Structures: one array with the elements for each entry.

For example when we want to save x,y,z coordinates. When using a Structure of Arrays we have one struct and three arrays for each coordinate.
If we use an Array of Structures we have a struct array with three entries. Each entry stores the values of the x, y, z coordinates.

A Structure of Arrays is good for vectorization. In our example with the coordinates, we can access all the entries of a specific coordinate
through the respective array.
With a SoA the arrays are kept seperate for each structure field. Memory access is contiguous if we perform vectorization of the structure instances.
When performing operations on one array of the structure, SoA leads better bandwidth usage.

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{8}
